---
title: "armadillo-illustration"
output: html_document
date: "2024-03-04"
editor_options:
  chunk_output_type: console
---

## Research question and data
Our overall research question is: which urinary biomarkers are associated with a risk of pancreatic cancer?

We use simulated data (source :https://www.kaggle.com/code/gkitchen/predicting-pancreatic-cancer). Here I've divided it as if
it comes from two datasets stored in (i) Groningen and (ii) Barcelona.

## Load required packages
```{r setup, include=FALSE}
library(DSMolgenisArmadillo)
library(dsBaseClient)
library(dsHelper)
library(tidyverse)
```

## Logging in
First, we need to log into two armadillo servers which store this data. First we get a token to log in. Normally we would need one for each server, but because we are using the demo server to host data from two cohorts we can use one here. A browser window should open, now click 'submit' (you can close the tab after):

```{r get-token}
url <- "https://armadillo-demo.molgenis.net/"
token <- armadillo.get_token(url)
```
Now we use this token to log in to two servers.

```{r make-login-object}
builder <- DSI::newDSLoginBuilder()

builder$append(
  server = "barcelona",
  url = url,
  token = token,
  table = "armadillo-illustration/barcelona/pancreatic",
  driver = "ArmadilloDriver",
  profile = "xenon")

builder$append(
  server = "groningen",
  url = url,
  token = token,
  table = "armadillo-illustration/groningen/pancreatic",
  driver = "ArmadilloDriver",
  profile = "xenon")

logindata <- builder$build()

conns <- DSI::datashield.login(logins = logindata, assign = TRUE, symbol = "pancreatic")
```
## Exploring data
Now we're logged in, let's see what we have. You can list all of the objects
available on the server using `ds.ls`():

```{r see-object}
ds.ls()
```

You can see that is one object here on each server, a dataset called "pancreatic"

Let's try to have a look at the data:
```{r}
ds.head("pancreatic")
```

This doesn't work, which is good! There is no datashield function called `ds.head`, as the whole idea of DataSHIELD is that you can't
see the data. What we can do instead is to ...

Inspect object class:
```{r dims}
ds.class("pancreatic")
```

Inspect object dimensions:
```{r dims}
ds.dim("pancreatic")
```

Get column names:
```{r cols}
ds.colnames("pancreatic")
```
## Summarising the data

We can also fetch summary statistics. To prevent disclosure of individual data,
as a default cell counts <3 are suppressed.
```{r summary}
exposures <- c("plasma_CA19_9", "creatinine", "LYVE1", "REG1B", "TFF1")
outcome <- "diagnosis"
covariates <- c("age", "sex")

descriptives <- dh.getStats(
  df = "pancreatic", 
  vars = c(exposures, outcome, covariates))
```

For example, we can see the number of cancer cases in each sample:
```{r exposures}
descriptives$categorical %>%
  dplyr::select(variable, cohort, category, value, perc_valid) %>%
  dplyr::filter(variable == "diagnosis" & cohort != "combined")
```
We can also view summaries of the biomarkers:

```{r outcomes}
descriptives$continuous %>%
  dplyr::select(variable, cohort, mean, std.dev) %>%
  dplyr::filter((variable %in% exposures) & cohort != "combined")
```

## Analysing the data
To see whether any of the biomarkers in our sample are associated with cancer 
diagnosis, we can use a logistic regression model. Here we use a function
to perform a two stage analysis:

1. Estimate the coefficients separately in each cohort,
2. Combine the estiamtes by meta-analysis

```{r}
reg_output <- ds.glmSLMA(
  formula = "diagnosis~1+plasma_CA19_9+creatinine+REG1B",
  family = "binomial", 
  dataName = "pancreatic")
```
## View results

First we can get the coefficients for each exposure, both separately and the
combined estimate:
```{R results}
coefficients <- dh.lmTab(
  model = reg_output, 
  type = "glm_slma", 
  coh_names = c("barcelona", "groningen"), 
  direction = "wide", 
  ci_format = "separate", 
  family = "binomial",
  exponentiate = T)

coefficients
```
We can also make a forest plot, which is more pleasant to look at. First clean
the output a little:

```{r preplot}
plot_data <- coefficients %>% 
  dplyr::mutate(cohort = case_when(
    cohort == "barcelona" ~ "Barcelona", 
    cohort == "groningen" ~ "Groningen", 
    cohort == "combined" ~ "Combined")) %>%
  mutate(cohort = factor(cohort, levels = c("Barcelona", "Groningen", "Combined"), 
                         ordered = T)) %>%
  dplyr::filter(variable != "intercept")
```

Now make the plot:

```{r}
source("forest_theme.R")

plot_data %>%
  ggplot(aes(y = variable)) +
  geom_point(aes(x=est), shape=15, size=3) +
  geom_linerange(aes(xmin=lowci, xmax=uppci)) +
  labs(x="Odds Ratio") +
  coord_cartesian(ylim=c(1,3), xlim=c(-5, 5)) +
  geom_vline(xintercept = 0, linetype="dashed") +
  annotate("text", x = -.2, y = 11, label = "Protective") +
  annotate("text", x = .2, y = 11, label = "Harmful") +
  facet_wrap(~cohort, ncol = 1) +
  forest_theme
```

## Summary
Whilst there is some evidence for an association between creatinine and risk
of pancreatic cancer in Groningen, the estimate is very imprecise. In the meta-
analysed results, we see little evidence that any of these biomarkers are associated
with risk of pancreatic cancer.


